Time Series Length at Which Block Bootstrapping is Effective for Estimation of Parameter and Its Variance

### 2022-12-08 Manuscript/manuscript.tex
Repo Update

In this experiment, the mean and auto-correlation function (with a maximum lag of 1) of a 1000 block-bootstrapped time series samples are computed, 
and a 95\% confidence intervals of these parameters is created. This is replicated 4000 times, 
and the proportion of confidence intervals that recover the true known parameter 
(coverage rate) is recorded. The goal of this is to observe how successful the bootstrapping process
is at estimating a parameter and its variance. The key variable being observed was, 
the length of the time series, or the size of the sample. It is known that as n
increases, block bootstrapping will become a less accurate method for estimation
(the coverage rates will decrease). The question is at what range of sample lengths does this
start to become a problem and what range of sample lengths is necessary to reach a coverage rate of 95\%.

@@

For all simulation studies, the mean can be kept at 0, but the AR(1) Coefficient is the main factor that makes block bootstrap less effective, so its impact on the method's performance should be analyzed. For a certain variation of the ARIMA process, a block bootstrap 95\% confidence interval for a sample of length 32 is replicated 4000 times (each replication of the method is applied to a newly generated ARIMA simulation). For each individual confidence interval, it can be recorded whether or not the interval includes the true parameter. The coverage rate, or the proportion of intervals that include the true parameter, can then be recorded. If the parameter and its variance are being properly estimated by the block bootstrap, the coverage rate should reflect this by being close to the confidence level of 95\%. If this is not the case, the method is not working as well as it should. The width of the confidence interval for each replication is dependent on the variance of the parameter of the time series. If the interval is on average not wide enough to capture the parameter at the expected rate of 95\%, it indicates that the variance of the parameter (which will be estimated from the standard error of the parameter estimates from the bootstrapped samples) will be underestimated. If the interval is too wide and captures the parameter at a rate above 95\%, it indicates that the variance of the parameter will be overestimated. Since the coverage rate is a proportion based on many replication of independent Bernoulli outcomes, a 95\% confidence interval of the coverage rate at that sample size can also be created. We can be 95\% confident that the true expected coverage rate is included in this interval, so if .95 is included in the interval, this indicates that the block bootstrap method is working well. Once the simulation is done for this sample size, the process can be run again with a sample twice as big. This can be continued until the method is determined to be working acceptably (.95 is included within the confidence interval of the coverage rate).

@@

12 variations of simulation studies were analyzed based on factors that might affect the results of the block bootstrap method. Simulation studies were conducted for an autoregressive process with AR = 0, an autoregressive process with AR = 0.2, and an autoregressive process with AR = 0.4. For each of these AR(1) coefficient values, separate simulation studies were conducted for the mean and AR(1) coefficient. For each of these variations, the non-moving block bootstrap method as well as the moving block bootstrap method were applied to estimate the respective parameter.

@@

As the AR(1) coefficient increases, it was found unsurprisingly that a large sample size was needed for the block bootstrap method to work. When the target for estimation was the mean of the process, moving block bootstrap generally had a similar performance to non-moving block bootstrap. However when the target for estimation was the AR(1) coefficient of the process, moving block bootstrap noticeably recovered the parameter at a greater rate than non-moving block bootstrap. When the target was the mean and theta was more than 0, the coverage rate was generally much higher than that for when the target was the AR(1) coefficient. However when theta was 0, the AR(1) coefficient was recovered at a greater rate than the mean.

@@

This study finds that as the AR coefficient increases, the sample size necessary for effective block bootstrap estimation must be larger. When implementing the block bootstrap method to estimate the mean of an autoregressive process, the non-moving method and moving method perform about the same. However, when estimating the AR(1) coefficient, the non-moving method performs noticeably worse than the moving method, which still covers the coefficient poorly when theta is greater than 0.

>>>>

In this experiment, the mean, auto-correlation function (with a maximum lag of 1), and variance of a 1000 block-bootstrapped time series samples are computed, 
and a 95\% confidence intervals of these parameters is created. This is replicated 4000 times, 
and the proportion of confidence intervals that recover the true known parameter 
(coverage rate) is recorded. The goal of this is to observe how successful the bootstrapping process
is at estimating a parameter and its variance. The key variable being observed was, 
the length of the time series, or the size of the sample. It is known that as n
increases, block bootstrapping will become a less accurate method for estimation
(the coverage rates will decrease). The question is at what range of sample lengths does this
start to become a problem and what range of sample lengths is necessary to reach a coverage rate of 95\%.

@@

For all simulation studies, the mean can be kept at 0, but the AR(1) Coefficient is the main factor that makes block bootstrap less effective, so its impact on the method's performance should be analyzed. For a certain variation of the ARIMA process, block bootstrap 95\% confidence intervals for each of 3 parameters (mean, AR(1) Coefficient, and sigma-squared) for a sample of length 32 is replicated 4000 times (each replication of the method is applied to a newly generated ARIMA simulation). For each individual confidence interval, it can be recorded whether or not the interval includes the true parameter. The coverage rate, or the proportion of intervals that include the true parameter, can then be recorded. If the parameter and its variance are being properly estimated by the block bootstrap, the coverage rate should reflect this by being close to the confidence level of 95\%. If this is not the case, the method is not working as well as it should. The width of the confidence interval for each replication is dependent on the variance of the parameter of the time series. If the interval is on average not wide enough to capture the parameter at the expected rate of 95\%, it indicates that the variance of the parameter (which will be estimated from the standard error of the parameter estimates from the bootstrapped samples) will be underestimated. If the interval is too wide and captures the parameter at a rate above 95\%, it indicates that the variance of the parameter will be overestimated. Since the coverage rate is a proportion based on many replication of independent Bernoulli outcomes, a 95\% confidence interval of the coverage rate at that sample size can also be created. We can be 95\% confident that the true expected coverage rate is included in this interval, so if .95 is included in the interval, this indicates that the block bootstrap method is working well. Once the simulation is done for this sample size, the process can be run again with a sample twice as big. This can be continued until the method is determined to be working acceptably (.95 is included within the confidence interval of the coverage rate).

@@

6 variations of simulation studies were analyzed based on factors that might affect the results of the block bootstrap method. Simulation studies were conducted for an autoregressive process with AR = 0, an autoregressive process with AR = 0.2, and an autoregressive process with AR = 0.4. For each of these AR(1) coefficient values, separate simulation studies were conducted for the mean and AR(1) coefficient. For each of these variations, the non-moving block bootstrap method as well as the moving block bootstrap method were applied to estimate the respective parameter.

@@

As the AR(1) coefficient increases, it was found unsurprisingly that a large sample size was needed for the block bootstrap method to work. When the target for estimation was the mean or the variance of the process, moving block bootstrap generally had a similar performance to non-moving block bootstrap. However when the target for estimation was the AR(1) coefficient of the process, moving block bootstrap noticeably recovered the parameter at a greater rate than non-moving block bootstrap. When the target was the mean or the variance and theta was more than 0, the coverage rate was generally much higher than that for when the target was the AR(1) coefficient. However when theta was 0, the AR(1) coefficient was recovered at a greater rate than the mean or variance .

@@

This study finds that as the AR coefficient increases, the sample size necessary for effective block bootstrap estimation must be larger. When implementing the block bootstrap method to estimate the mean or variance of an autoregressive process, the non-moving method and moving method perform about the same. However, when estimating the AR(1) coefficient, the non-moving method performs noticeably worse than the moving method, which still covers the coefficient poorly when theta is greater than 0. Whereas the performance improves as sample length increases when the target is the mean or variance, the performance decays as sample length increases when the target is the AR(1) coefficient.
